---
title: 'Dehumanization X Columbus Day/Indigenous Peoples Day'
author: "Terrence Pope"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    theme: spacelab
    highlight: zenburn
---

```{r setup,include=F}
knitr::opts_chunk$set(echo = T)
library(haven)
library(tidyverse)
library(magrittr)
library(pander)
library(psych)
library(Hmisc)
library(reshape2)

#Creating a function that helps summarize data
summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE,
                      conf.interval=.95, .drop=TRUE) {
  library(plyr)
  
  # New version of length which can handle NA's: if na.rm==T, don't count them
  length2 <- function (x, na.rm=FALSE) {
    if (na.rm) sum(!is.na(x))
    else       length(x)
  }
  
  # This does the summary. For each group's data frame, return a vector with
  # N, mean, and sd
  datac <- plyr::ddply(data, groupvars, .drop=.drop,
                       .fun = function(xx, col) {
                         c(N    = length2(xx[[col]], na.rm=na.rm),
                           mean = mean   (xx[[col]], na.rm=na.rm),
                           sd   = sd     (xx[[col]], na.rm=na.rm)
                         )
                       },
                       measurevar
  )
  
  # Rename the "mean" column    
  datac <- plyr::rename(datac, c("mean" = measurevar))
  
  datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean
  
  # Confidence interval multiplier for standard error
  # Calculate t-statistic for confidence interval: 
  # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
  ciMult <- qt(conf.interval/2 + .5, datac$N-1)
  datac$ci <- datac$se * ciMult
  
  return(datac)
}

#load in data
path <- file.path("~/Desktop/UW/Summer '18/CultureColab/Native Bias/Data/Study_2_w_composites.sav")

studytwodata <- read_sav(path)

dfdehumanDV<- as.data.frame(cbind(studytwodata$Dehuman_wht_1,studytwodata$Dehuman_wht_2,studytwodata$Dehuman_wht_3,studytwodata$Dehuman_wht_4,studytwodata$Dehuman_wht_5,studytwodata$Dehuman_wht_6,studytwodata$Dehuman_wht_7,studytwodata$Dehuman_wht_8,studytwodata$Dehuman_wht_9,studytwodata$Dehuman_wht_10,studytwodata$Dehuman_wht_11,studytwodata$Dehuman_wht_12,studytwodata$Dehuman_wht_13,studytwodata$Dehuman_wht_14,studytwodata$Dehuman_wht_15,studytwodata$Dehuman_wht_16,studytwodata$Dehuman_wht_17,studytwodata$Dehuman_wht_18,studytwodata$Dehuman_wht_19,studytwodata$Dehuman_wht_20,studytwodata$Dehuman_Native_1,studytwodata$Dehuman_Native_2,studytwodata$Dehuman_Native_3,studytwodata$Dehuman_Native_4,studytwodata$Dehuman_Native_5,studytwodata$Dehuman_Native_6,studytwodata$Dehuman_Native_7,studytwodata$Dehuman_Native_8,studytwodata$Dehuman_Native_9,studytwodata$Dehuman_Native_10,studytwodata$Dehuman_Native_11,studytwodata$Dehuman_Native_12,studytwodata$Dehuman_Native_13,studytwodata$Dehuman_Native_14,studytwodata$Dehuman_Native_15,studytwodata$Dehuman_Native_16,studytwodata$Dehuman_Native_17,studytwodata$Dehuman_Native_18,studytwodata$Dehuman_Native_19,studytwodata$Dehuman_Native_20,studytwodata$DV1_4,studytwodata$DV1_5,

studytwodata$BorninUS,studytwodata$LengthState,studytwodata$FatherUSBorn,studytwodata$MotherUSBorn,studytwodata$PrimaryCaretaker,studytwodata$PrimaryEd,studytwodata$SecondCaretaker,studytwodata$SecondEdu,studytwodata$Religion,studytwodata$Religiosity,studytwodata$College,studytwodata$Year,studytwodata$Education,studytwodata$LibCon,studytwodata$SES,studytwodata$Race,studytwodata$StateNumeric,studytwodata$LongestRegion,studytwodata$SubjectNumber))




colnames(dfdehumanDV) <- c(V1 = "DH1_W_Responsible",V2 ="DH2_W_Civilized",V3 = "DH3_W_Moral",V4 = "DH4_W_Polite",
                           V5 = "DH5_W_Childlike",V6 = "DH6_W_Rational",V7 = "DH7_W_Warm",V8 = "DH8_W_Agentic",
                           V9 = "DH9_W_Refined",V10 = "DH10_W_Lacking_Culture",V11 = "DH11_W_Lacking_Self-restraint",V12 = "DH12_W_Instinctual",
                           V13 = "DH13_W_Mature",V14 = "DH14_W_Stoic",V15 = "DH15_W_Emotionally_Responsive",V16 = "DH16_W_Cold",
                           V17 = "DH17_W_Open",V18="DH18_W_Rigid",V19 =  "DH19_W_Passive",V20 = "DH20_W_Superficial",
                           V21 =  "DH1_N_Responsible",V22 = "DH2_N_Civilized",V23 = "DH3_N_Moral",V24 = "DH4_N_Polite",
                           V25 = "DH5_N_Childlike",V26 = "DH6_N_Rational",V27 = "DH7_N_Warm",V28 = "DH8_N_Agentic",
                           V29 =  "DH9_N_Refined",V30 = "DH10_N_Lacking_Culture",V31 = "DH11_N_Lacking_Self-restraint",V32 =  "DH12_N_Instinctual",
                           V33 = "DH13_N_Mature",V34 ="DH14_N_Stoic",V35 = "DH15_N_Emotionally_Responsive",V36 = "DH16_N_Cold",
                           V37 = "DH17_N_Open",V38 = "DH18_N_Rigid",V39 = "DH19_N_Passive",V40 =  "DH20_N_Superficial", 
                           V41 = "No_Columbus",V42 = "Create_IPD",V44 = "BorninUS",V45 = "LengthState",V46 = "FatherUSBorn", 
                           V48 = "MotherUSBorn",V49 = "PrimaryCaretaker",V50 = "PrimaryEd",V51 = "SecondCaretaker",V52 = "SecondEdu",
                           V53 = "Religion",V54 = "Religiosity",V55 = "College",V56 = "Year",V57 = "Education",V58 = "LibCon",V59 = "SES",
                           V60 = "Race",V61 = "StateNumeric",V62 = "LongestRegion",V63 = "SubjectNumber")

dfdehumanDV <- dfdehumanDV %>% 
  mutate(White_Dehumanize = rowMeans(dfdehumanDV[,1:20]))
dfdehumanDV <- dfdehumanDV %>% 
  mutate(Native_Dehumanize = rowMeans(dfdehumanDV[,21:40]))
dfdehumanDV <- dfdehumanDV %>%
  mutate(Age = as.numeric(studytwodata$Age))

dfdehumanDV$BorninUS <- factor(dfdehumanDV$BorninUS,
                               levels = c(0,1),
                               labels = c("No", "Yes"))
dfdehumanDV$FatherUSBorn <- factor(dfdehumanDV$FatherUSBorn,
                               levels = c(0,1,3),
                               labels = c("No", "Yes", "Unsure/Not applicable"))

dfdehumanDV$MotherUSBorn <- factor(dfdehumanDV$MotherUSBorn,
                               levels = c(0,1,3),
                               labels = c("No", "Yes", "Unsure/Not applicable"))

dfdehumanDV$PrimaryCaretaker <- factor(dfdehumanDV$PrimaryCaretaker,
                               levels = c(1,2,3,4,5,6,7),
                               labels = c("Mother", "Father", "Extended female family member (e.g., grandmother, aunt)", "Extended male family member (e.g., grandfather, uncle)", "Legal guardian female", "Legal guardian male", "Another caretaker (please describe):"))

dfdehumanDV$SecondCaretaker <- factor(dfdehumanDV$SecondCaretaker,
                               levels = c(1,2,3,4,5,6,7,8),
                               labels = c("No, I did not have another caretaker", "Mother", "Father", "Extended female family member (e.g., grandmother, aunt)", "Extended male family member (e.g., grandfather, uncle)", "Legal guardian female", "Legal guardian male", "Another caretaker (please describe):"))

dfdehumanDV$Religion <- factor(dfdehumanDV$Religion,
                               levels = c(1,2,3,4,5,6,7,8),
                               labels = c("None", "Catholic", "Christian (non Catholic)", "Jewish", "Muslim", "Buddhist", "Hindu", "Other (please specify):"))

dfdehumanDV$College <- factor(dfdehumanDV$College,
                               levels = c(1,2),
                               labels = c("Yes", "No"))

dfdehumanDV$SES <- factor(dfdehumanDV$SES,
                               levels = c(-1,1),
                               labels = c("No College", "College"))

dfdehumanDV$Race <- factor(dfdehumanDV$Race,
                               levels = c(1,2,3,4,5,6,7,8),
                               labels = c("Asian", "Black","White","Latino","Middle Eastern","Native","Other","Multiracial"))

dfdehumanDV$StateNumeric <- factor(dfdehumanDV$StateNumeric,
                               levels = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24,25,26,27,28,29,30,31,32,33,34,35,36,37,38,39,40,41,42,43,44,45,46,47,48,49,50),
                               labels = c("Alabama","Alaska","Arizona","Arkansas","California","Colorado","Connecticut","Delaware","Florida","Georgia","Hawaii","Idaho","Illinois","Indiana","Iowa","Kansas","Kentucky","Louisiana","Maine","Maryland","Massachusetts","Michigan","Minnesota","Mississippi","Missouri","Montana","Nebraska","Nevada","New Hampshire","New Jersey","New Mexico","New York","North Carolina","North Dakota","Ohio","Oklahoma","Oregon","Pennsylvania","Rhode Island","South Carolina","South Dakota","Tennessee","Texas","Utah","Vermont","Virginia","Washington","West Virginia","Wisconsin","Wyoming"))

dfdehumanDV$LongestRegion <- factor(dfdehumanDV$LongestRegion,
                               levels = c(1,2,3,4,5,6,7,8),
                               labels = c("New England","Mideast","Great Lakes","Plains","Southeast","Southwest","Rocky Mountain","Far West"))

destable <- as.data.frame(psych::describe(dfdehumanDV))
```
#Intro 

This is a data set that includes the following variables:

```{r names, include =T, echo=F}
variable.list <- as.matrix(names(dfdehumanDV))
pander(variable.list)
```
<br></br>
It's primary focus is to look a bit more closely at the dehumaniztion variables and how they interact not only with one another, but with the dependent variables (Columbus Day/Indigenous Peoples' Day (henceforth referred to as IPD)), as well as some of the demographic information gathered on the participants in this sample.

This sample comes from **Study Two**, so it is (*if I'm not mistaken*) a sample of participants gathered via *mturk*.

#Descriptives
First, lets take a look at a descriptives table of our variables. Keep in mind that some of the variables are categorical and are more difficult to interpret in a descriptives table.

```{r descriptives table, include =T, echo=F}
destable <- as.data.frame(cbind(destable[,2:5],destable[,13]))
table.names <- c("n", "mean", "sd", "median","se")
names(destable) <- table.names
pander(destable)
```

##Correlation of Dehumanization
One thing you'll probably notice right away is that I've made some aggregate variables at the end of the data set that average across all Dehumanization measures for both whites and natives. Let's take a look at the relationship between the two zoomed out a bit.
```{r whitenativedhcor, include =T, echo=F}
plot(dfdehumanDV$White_Dehumanize, dfdehumanDV$Native_Dehumanize)

dehum.nw.cor <- cor.test(dfdehumanDV$White_Dehumanize, dfdehumanDV$Native_Dehumanize, method = ("pearson"))

dehum.nw.cor
```

At baseline the two seem to hang together pretty well. People seem to be rating Natives and Whites similarly among the measures of dehumanization that we have.

##Dependent Variables: Columbus Day X Dehumanization
Before we move on, let's take a gander at the dependent variables. Does there seem to be a relationship between dehumanization and *Support for getting rid of Columbus Day*? 

```{r dv4cor, include = T, echo = F}
dv4cor <- dfdehumanDV %>%
  dplyr::select(SubjectNumber, No_Columbus, White_Dehumanize, Native_Dehumanize)

dv4cor.lng <- melt(dv4cor, id=c("SubjectNumber", "No_Columbus"), variable.name = "Group", value.name="Rating")

ggplot(dv4cor.lng, aes(x= Rating, y = No_Columbus, color = Group, group = Group)) +
  geom_point(position = position_jitter(width = 0.25)) +
  geom_smooth(stat = "smooth", position = "identity", method = "lm", se = T, na.rm = T) +
  xlab("Dehumanization (0-100)") +
  ylab("Support for No Columbus Day") +
  labs(color='Dehumanized Group') +
  scale_y_continuous(limits = c(0, NA))

lm1 <- lm(No_Columbus ~ Rating + Group + Rating:Group, dv4cor.lng)
summary(lm1)
```
`**Keep in mind while looking at this graph that low scores on the dehumanization measure mean that participants are dehumanizing the group` **more**.

<br></br>

##Dependent Variables: Indigenous Peoples' Day X Dehumanization
What about between dehumanization and *Support for IPD*?

```{r dv5cor, include = T, echo = F}
dv5cor <- dfdehumanDV %>%
  dplyr::select(SubjectNumber, Create_IPD, White_Dehumanize, Native_Dehumanize)

dv5cor.lng <- melt(dv5cor, id=c("SubjectNumber", "Create_IPD"), variable.name = "Group", value.name="Rating")

ggplot(dv5cor.lng, aes(x= Rating, y = Create_IPD, color = Group, group = Group)) +
  geom_point(position = position_jitter(width = 0.25)) +
  geom_smooth(stat = "smooth", position = "identity", method = "lm", se = T, na.rm = T) +
  xlab("Dehumanization (0-100)") +
  ylab("Support for IPD") +
  labs(color='Dehumanized Group') +
  scale_y_continuous(limits = c(0, NA))

lm2 <- lm(Create_IPD ~ Rating + Group + Rating:Group, dv5cor.lng)
summary(lm2)
```

##Dependent Variables: Side by Sides
We can also look at both of these groups side by side:

```{r, include = T, echo = F}
ggplot(dv4cor.lng, aes(x= Rating, y = No_Columbus, color = Group, group = Group)) +
  geom_point(position = position_jitter(width = 0.25)) +
  geom_smooth(stat = "smooth", position = "identity", method = "lm", se = T, na.rm = T) +
  xlab("Dehumanization (0-100)") +
  ylab("Support for No Columbus Day") +
  labs(color='Dehumanized Group') +
  facet_wrap(~ Group) +
  scale_y_continuous(limits = c(0, NA))

```
<br></br>
and...
<br></br>
```{r, include = T, echo = F}
ggplot(dv5cor.lng, aes(x= Rating, y = Create_IPD, color = Group, group = Group)) +
  geom_point(position = position_jitter(width = 0.25)) +
  geom_smooth(stat = "smooth", position = "identity", method = "lm", se = T, na.rm = T) +
  xlab("Dehumanization (0-100)") +
  ylab("Support for IPD") +
  labs(color='Dehumanized Group') +
  facet_wrap(~ Group) +
  scale_y_continuous(limits = c(0, NA))
```

##Dependent Variables: Digging a Little Deeper
Whe can see more clearly that significant interaction when they are side by side. Attitudes toward white more strongly predicts the relationship between dehumanization and support for removing Columbus day.

Interestingly, the direction of that relationship changes when we're talking about IPD. If we create a model that looks at this interaction with both ratings of support and dehumanization centered at the mean we see that this change in direction is significant (*p* = .005)

```{r, include = T, echo = F}
dv4and5cor <- dfdehumanDV %>%
  dplyr::select(SubjectNumber, Create_IPD, No_Columbus, Native_Dehumanize)

dv4and5cor.lng <- melt(dv4and5cor, id=c("SubjectNumber", "Native_Dehumanize"), variable.name = "Holiday_Support", value.name="Rating")

dv4and5cor.lng.cen <- dv4and5cor.lng
dv4and5cor.lng.cen$Native_Dehumanize <- scale(dv4and5cor.lng.cen$Native_Dehumanize, scale = F)
dv4and5cor.lng.cen$Rating <- scale(dv4and5cor.lng.cen$Rating, scale = F)

lm3 <- lm(Native_Dehumanize ~ Rating + Holiday_Support + Rating:Holiday_Support, dv4and5cor.lng.cen)
summary(lm3)
```
`**For some reason in the markdown document the "Rating" coefficient is significant. In the code it is not` (*p*=.12).

#Exploratory Analysis

Let's dig into some of the demographic information in this subset of the data.

##Race
First let's look at the frequency of our different racial groups
```{r race1, include = T, echo = F}
Race.obj <- describe(dfdehumanDV$Race)
Race.obj <- Race.obj$values
Race.bars <- as.vector(Race.obj$frequency)
names(Race.bars) <- Race.obj$value
pander(Race.bars)

colors <- c("red", "yellow", "green", "violet", "orange", "blue", "pink", "cyan") 
barplot(Race.bars, col=colors)
```

Since our data skews toward **Asian**, **Black**, **White**, and **Latino** let's use these groups in future analysis. 

###Race X Support for Abolishing Columbus Day
First off, let's look at how our different racial groups answered the question about Columbus Day.

```{r bardv1, include = T, echo = F}

dfdehumanDV.race <- 


# ggplot(Data.lng.c2, aes(x=Condition, y=Rating, fill=Race)) +
#   geom_bar(position=position_dodge(), stat="identity",
#            colour="black", # Use black outlines,
#            size=.3) +      # Thinner lines
#   geom_errorbar(aes(ymin=Rating-se, ymax=Rating+se),
#                 size=.3,    # Thinner lines
#                 width=.2,
#                 position=position_dodge(.9)) +
#   xlab("Participant Condition") +
#   ylab("Participant Ratings (1-7)") +
#   scale_fill_hue(name="Applicant Race", # Legend label, use darker colors
#                  breaks=c("AF", "AA"),
#                  labels=c("African American", "Asian American")) +
#   ggtitle("Ratings of Applicants on Measures of\nAmericanness, Qualification, and Status") +
#   scale_y_continuous(limits = c(0,7)) +
#   facet_wrap(~Measure) +
#   theme_bw()

# ggplot(data = gapminder, aes(x = year, y = lifeExp, group = country)) +
#   geom_line(alpha = 0.5, aes(color = "Country", size = "Country")) +
#   geom_line(stat = "smooth", method = "loess",
#             aes(group = continent, color = "Continent",
#                 size = "Continent"), alpha = 0.5) +
#   xlab("Year") +
#   ylab("Life Expectancy") +
#   facet_wrap(~ continent, nrow = 2) +
#   scale_color_manual(name = "Unit", values = c("Country" = "black",
#                                                "Continent" = "blue")) +             scale_size_manual(name = "Unit", values = c("Country" = 0.5,
#                                               "Continent" = 3)) +
#   theme_minimal(base_size = 14) +
#   theme(legend.position=c(0.75, 0.2),
#         axis.text.x = element_text(angle = 45))
```
